Interpolation of Sparse High-Dimensional Data\thanks {This work was supported by the National Science Foundation Grant CNS-1565314.}

### 2020-08-13 [2019-02]_NUMA_(interpolant_error_bound)/tchlux-NA-interpolation/Versions/2-paper.tex
Finished NUMA paper and submitted. Might need to do one more round with editors.

%% Include the response to reviewers at the beginning of the document.
\include{tchlux-NA-response-2020-04}

@@

The theoretical foundation of low dimensional interpolation allows the
construction of strong error bounds that are \changed{lacking
  in}{absent from} high dimensional problems. \changed{A large body of
  literature exists studying the construction and application of
  sparse grids \cite{bungartz2004sparse,chkifa2014high}, which attempt
  to address the problem of dimensional scaling. The sparse grid point
  sets combined with specific approximation functions are used to
  construct total error bounds by considering projections onto
  orthogonal polynomials
  \cite{migliorati2013approximation,nobile2016convergence}. However,
  sparse grids have exponential size in dimension making them
  intractable for more than tens of dimensions and do not immediately
  apply to the problem of scattered data interpolation.}{} This work
extends some known results regarding the secant method
\cite{dennis1996numerical} to construct \changed{a uniform error bound
  for interpolants over arbitrary scattered data in any dimension}{an
  interpolation error bound for problems of arbitrary
  dimension}. These error bounds are useful, considering
\changed{similar uniform bounds}{the same cannot} cannot be
\changed{constructed}{said} for regression algorithms in general. The
maximum complexity of an interpolant is bounded by the amount of data
available, while the maximum complexity of a regressor is bounded by
both the amount of data and the chosen parametric form. For this
reason, generic uniform bounds are largely unobtainable for regression
techniques on arbitrary approximation problems, even when the
approximation domain is bounded.

@@

\subsection{Multilayer Perceptron Regressor}
\label{sec:mlp}
The neural network is a well studied and widely used method for both
regression and classification tasks
\cite{hornik1989multilayer,rumelhart1988learning}. When using the
rectified linear unit (ReLU) activation function
\cite{dahl2013improving} and fitting the model with stochastic
gradient descent (SGD) or BFGS minimization techniques
\cite{goh2017why,moller1993scaled,robbins1951stochastic}, the model
built by a multilayer perceptron uses layers $l : \mathbb{R}^{i}
\rightarrow \mathbb{R}^{j}$ defined by
 $$ l(u) = \bigl( u^t W_l + c_l \bigr)_+ ,$$
where $c_l \in \mathbb{R}^j$ and $W_l$ is the $i \times j$ weight
matrix for layer $l$. In this form, the multilayer perceptron (MLP)
produces a piecewise linear model of the input data. The computational
complexity of training a multilayer perceptron is $\mathcal{O}(n d
m)$, where $m$ is determined by the sizes of the layers of the network
and the stopping criterion of the minimization used for finding
weights. This paper uses an MLP built from Keras and Tensorflow to
perform regression \cite{tensorflow2015-whitepaper,chollet2015keras}
with ten hidden layers each having thirty two nodes (a total of
approximately ten thousand parameters), ReLU activation, and
\changed{one hundred}{five} thousand \changed{steps}{epochs} of SGD
for training. \changed{It should be noted that the choice of neural
  network architecture impacts approximation performance, but no
  architecture tuning is performed here}{}.

@@

The Delaunay triangulation is a well-studied geometric technique for
producing an interpolant \cite{lee1980two}. The Delaunay triangulation
of a set of data points into simplices is such that there are no data
points inside the sphere defined by the vertices of each simplex. For
a $d$-simplex S with vertices $x^{(0)}$, $x^{(1)}$, $\ldots$,
$x^{(d)}$, and function values $f\bigl(x^{(i)}\bigr)$, $i=0$,
$\ldots$, $d$, $y \in S$ is a unique convex combination of the
vertices,
 $$ y = \sum_{i=0}^{d} w_i x^{(i)}, \quad \sum_{i=0}^{d} w_i = 1, \quad w_i \geq 0, \quad i=0,\ldots,d, $$
and the Delaunay interpolant $\hat f(y)$ at $y$ is given by
 $$ \hat f(y) = \sum_{i=0}^{d} w_i f\bigl(x^{(i)}\bigr). $$ The
computational complexity of Delaunay interpolation (for the
implementation used \cite{chang2018polynomial}) is $\mathcal{O}(k n
d^3)$. Given pathological data the entire triangulation could be
computed with $k = n^{\lceil d / 2 \rceil}$, but the average case yields
$k = d \log d$ and is capable of scaling reasonably to $d \leq 50$. In
the present application, a Delaunay simplex $S$ containing $y$ is
found, then the $d+1$ vertices (points in $X$) of $S$ are used to
assign weights to each vertex and produce the predicted function value
for point $y$.

@@

The modified Shepard method used here (ShepMod) generates a continuous
approximation based on Euclidean distance and resembles a nearest
neighbor interpolant \cite{cover1967nearest}. This model is a type of
\textit{metric interpolation}, also called a Shepard method
\cite{gordon1978shepard,shepard1968two}. The interpolant has the form
 $$ p(x) = \frac{\sum\limits_{k=1}^{n}W_k(x)f\bigl(x^{(k)}\bigr)}
     {\sum\limits_{k=1}^{n}W_k(x)} ,$$
where \changed{$W_k(x^{(k)}) = 1$, $W_k(x) = \bigl(\bigl(r_k -
  \bigl\|x - x^{(k)}\bigr\|_2\bigr)_+ \big/ \bigl( r_k \bigl\|x -
  x^{(k)}\bigr\|_2 \bigr)\bigr)^2$ for $x \neq x^{(k)}$,}{} and $r_k
\in \mathbb{R}$ is the smallest radius such that at least $d+1$ other
points $x^{(j)}$, $j \not = k$, are inside the closed Euclidean ball
of radius $r_k$ about $x^{(k)}$. The computational complexity of
ShepMod is $\mathcal{O}(n^2d)$. This paper uses a Fortran 95
implementation of ShepMod derived from SHEPPACK
\cite{thacker2010algorithm}.

\changed{
\begin{figure}
  \centering
  \includegraphics[width=0.9\textwidth,]{example_shepard_with_weight.pdf}
  \caption{\color{red} On the left above is a depiction of the radius
    of influence for three chosen points of a collection in two
    dimensions using the modified Shepard criteria. On the right is
    the shape of the weight function $W_k$ for all data points, where
    $W_k(x) = 0$ when $\|x-x^{(k)}\|_2 \geq r_k$. It can also be seen
    that in the limit as $W_k$ tends towards infinity at $x_k$, the
    relative weight of $x_k$ tends to one.
  \vspace{-.1cm}}
  \label{fig:shepard}
\end{figure}
}{}

@@

  Suppose $g'(0) > \g/2$. Then $|g'(0)-g'(t)| \le \g t \implies g'(0)
  - \g t \le g'(t)$, and the line $w=g'(0)-\g t$ intersects the
  $t$-axis at $\tilde t=g'(0)/\g > 1/2$. $\tilde t<1$ necessarily,
  since by Rolle's Theorem there exists $0<z<1$ such that $g'(z)=0$
  and so $g'(0) - \g z \le g'(z) = 0$. Now \changed{by integrating}{}
  $$\frac{g'(0)^2}{2\g} = \frac{g'(0)\tilde t}{2} = \int_0^{\tilde t}
  (g'(0)-\g t)dt \le \int_0^{\tilde t} g'(t)dt = g(\tilde t),$$ and
  \changed{acknowledging $g'(0) > \g / 2$}{}
  \begin{align*}
    \frac{-g'(0)^2}{2\g} < g'(0) - \frac{g'(0)^2}{2\g} - \g/2 =
    \frac{(1-\tilde t)(g'(0)-\g)}{2} &= \int_{\tilde t}^1 (g'(0)-\g
    t)dt \\ &\le \int_{\tilde t}^1 g'(t)dt = -g(\tilde t),
  \end{align*}
  \noindent a contradiction for the value of $g(\tilde t)$. A similar
  contradiction arises for $g'(0)<-\g/2$. Therefore $|g'(0)| \le
  \g/2$. In terms of $f$,
  $$\bigl|\langle \nabla f(x_0)-c,x_j-x_0\rangle\bigr| = |g'(0)| \le
  \gamma \|x_j-x_0\|_2^2/2 \le \gamma k^2 / 2,$$ which holds for all
  $1\le j\le d$.  Finally, using {\it Lemma \ref{lemma:2}},
  $$\|\nabla f(x_0)-c\|_2^2 \le \frac{d}{\sigma_d^2}(\gamma k^2/2)^2
  \implies \|\nabla f(x_0)-c\|_2 \le \frac{\sqrt{d} \, \gamma
    k^2}{2\sigma_d}.$$ \qed
\end{proofdot}

@@

In addition to the figures displaying approximation results for each
data set, tables of accompanying numerical results \changed{including
  timings and exact prediction errors}{} are located in the Appendix
(Section \ref{sec:appendix}). All of the test data sets capture
underlying functions that are almost certainly stochastic. As
described in Section \ref{sec:introduction}, regression techniques
appear most appropriate for these problems. However, typically data
grows exponentially more sparse with increasing dimension. Given that
sparse data regressors tend towards interpolation and, as demonstrated
in Section \ref{sec:analytic}, interpolants produce similar (if not
identical) results, it is presumed that interpolants are equally
viable approximation techniques on these problems.

@@

\bibliographystyle{spmpsci}      % mathematics and physical sciences
\bibliography{paper}

@@

\begin{table}[H]
  \centering
  \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
    \hline
    Algorithm & \multicolumn{2}{c}{Min}       & \multicolumn{2}{c}{$25^{th}$} & \multicolumn{2}{c}{$50^{th}$} & \multicolumn{2}{c}{$75^{th}$} & \multicolumn{2}{c|}{Max}\\
    \hline
    MARS      & $0$&$00984$                   & $3$&$11$                      & $7$&$01$                      & $\mathit{11}$&$\mathit{7}$    & $1090$&$0$\\
    SVR       & $0$&$0118$                    & $\mathbf{0}$&$\mathbf{615}$   & $\mathbf{0}$&$\mathbf{931}$   & $\mathbf{5}$&$\mathbf{89}$    & $1090$&$0$\\
    MLP       & $0$&$0426$                    & $2$&$63$                      & $5$&$27$                      & $14$&$0$                      & $1090$&$0$\\
    Delaunay  & $\mathbf{0}$&$\mathbf{0}$     & $1$&$98$                      & $5$&$37$                      & $13$&$1$                      & $\mathit{1080}$&$\mathit{0}$\\
    ShepMod   & $\mathbf{0}$&$\mathbf{0}$     & $1$&$93$                      & $6$&$27$                      & $16$&$0$                      & $1090$&$0$\\
    LSHEP     & $0$&$04$                      & $2$&$17$                      & $8$&$87$                      & $19$&$1$                      & $\mathbf{1070}$&$\mathbf{0}$\\
    Voronoi   & $\mathit{0}$&$\mathit{00982}$ & $3$&$65$                      & $7$&$56$                      & $15$&$6$                      & $1090$&$0$\\
    BoxSpline & $\mathbf{0}$&$\mathbf{0}$     & $\mathit{1}$&$\mathit{27}$    & $\mathit{4}$&$\mathit{61}$    & $12$&$6$                      & $1090$&$0$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-forest-fire}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum absolute errors respectively. The
    minimum of each column is boldface, while the second lowest value
    is italicized. All values are rounded to three significant
    digits.}
  \label{table:error-forest-fire}
\end{table}

\begin{table}[H]
  \centering
  \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
    \cline{1-3}\cline{5-10}
    Algorithm & \multicolumn{2}{c|}{\% Best}   &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)} & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
    \cline{1-3}\cline{5-10}
    MARS      & \quad$7$&$3$                   &  & \quad\quad$29$&$1$                      & \quad$0$&$00137$                   & \quad\quad\quad$0$&$0686$\\
    SVR       & $\mathbf{78}$&$\mathbf{0}$     &  & $\mathbf{0}$&$\mathbf{00584}$           & $\mathit{0}$&$\mathit{0000620}$    & $\mathit{0}$&$\mathit{00310}$\\
    MLP       & $0$&$0$                        &  & $32$&$8$                                & $0$&$000871$                       & $0$&$0436$\\
    Delaunay  & $0$&$2$                        &  & $0$&$0151$                              & $0$&$0234$                         & $1$&$18$\\
    ShepMod   & $2$&$0$                        &  & $\mathit{0}$&$\mathit{00634}$           & $0$&$0000644$                      & $0$&$00322$\\
    LSHEP     & $5$&$1$                        &  & $0$&$0275$                              & $\mathbf{0}$&$\mathbf{0000463}$    & $\mathbf{0}$&$\mathbf{00231}$\\
    Voronoi   & $0$&$0$                        &  & $0$&$0152$                              & $0$&$000396$                       & $0$&$0198$\\
    BoxSpline & $\mathit{9}$&$\mathit{7}$      &  & $0$&$00724$                             & $0$&$0000978$                      & $0$&$00489$\\
    \cline{1-3}\cline{5-10}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating forest fire data in Table
    \ref{table:error-forest-fire}. On the right columns are median fit
    time of 454 points, median time for one approximation, and median
    time approximating 50 points.}
  \label{table:best-forest-fire}
\end{table}

@@

\begin{table}
  \centering
  \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
    \hline
    Algorithm & \multicolumn{2}{c}{Min}                   & \multicolumn{2}{c}{$25^{th}$} & \multicolumn{2}{c}{$50^{th}$} & \multicolumn{2}{c}{$75^{th}$} & \multicolumn{2}{c|}{Max}\\
    \hline
    MARS      & $0$&$00948$                               & $9$&$98$                      & $20$&$4$                      & $32$&$9$                      & $863$&$0$\\
    SVR       & $0$&$00233$                               & $3$&$15$                      & $7$&$21$                      & $11$&$3$                      & $\mathbf{28}$&$\mathbf{6}$\\
    MLP       & $0$&$0000239$                             & $\mathit{0}$&$\mathit{533}$   & $\mathit{1}$&$\mathit{25}$    & $\mathbf{2}$&$\mathbf{84}$    & $39$&$3$\\
    Delaunay  & $\mathit{3}$&$\mathit{72 \times 10^{-12}}$ & $1$&$2$                       & $3$&$5$                       & $7$&$67$                      & $30$&$7$\\
    ShepMod   & $\mathbf{0}$&$\mathbf{0}$                 & $\mathbf{0}$&$\mathbf{255}$   & $\mathbf{0}$&$\mathbf{908}$   & $\mathit{3}$&$\mathit{43}$    & $34$&$5$\\
    LSHEP     & $0$&$00254$                               & $2$&$93$                      & $7$&$16$                      & $13$&$1$                      & $\mathit{29}$&$\mathit{0}$\\
    Voronoi   & $\mathbf{0}$&$\mathbf{0}$                 & $1$&$29$                      & $3$&$52$                      & $6$&$87$                      & $30$&$1$\\
    BoxSpline & $0$&$006$                                 & $4$&$3$                       & $8$&$91$                      & $15$&$1$                      & $45$&$3$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-parkinsons}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum absolute errors respectively. The
    minimum of each column is boldface, while the second lowest value
    is italicized. All values are rounded to three significant
    digits.}
  \label{table:error-parkinsons}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
    \cline{1-3}\cline{5-10}
    Algorithm & \multicolumn{2}{c|}{\% Best} &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)} & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
    \cline{1-3}\cline{5-10}
    MARS      & \quad\,$0$&$0$               &  & \quad\quad\quad$37$&$9$                 & \quad\,\,$0$&$00253$               & \quad\quad\quad\,\,$1$&$48$\\
    SVR       & $0$&$1$                      &  & $\mathbf{0}$&$\mathbf{859}$             & $\mathit{0}$&$\mathit{000181}$     & $\mathit{0}$&$\mathit{106}$\\
    MLP       & $\mathit{32}$&$\mathit{0}$   &  & $348$&$0$                               & $0$&$00111$                        & $0$&$653$\\
    Delaunay  & $0$&$0$                      &  & $2$&$47$                                & $1$&$22$                           & $720$&$0$\\
    ShepMod   & $\mathbf{66}$&$\mathbf{4}$   &  & $\mathit{1}$&$\mathit{13}$              & $0$&$000182$                       & $0$&$107$\\
    LSHEP     & $0$&$0$                      &  & $2$&$39$                                & $\mathbf{0}$&$\mathbf{000144}$     & $\mathbf{0}$&$\mathbf{0845}$\\
    Voronoi   & $1$&$6$                      &  & $2$&$77$                                & $0$&$0274$                         & $16$&$1$\\
    BoxSpline & $0$&$0$                      &  & $1$&$26$                                & $0$&$000643$                       & $0$&$377$\\
    \cline{1-3}\cline{5-10}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating Parkinson's data in Table
    \ref{table:error-parkinsons}. On the right columns are median fit
    time of 5288 points, median time for one approximation, and median
    time approximating 587 points.}
  \label{table:best-parkinsons}
\end{table}

>>>>

%% Include the response to reviewers at the beginning of the document.
\include{tchlux-NA-response-2_2020-04}

@@

The theoretical foundation of low dimensional interpolation allows the
construction of strong error bounds that are \changed{lacking
  in}{absent from} high dimensional problems. \changed{A large body of
  literature exists studying the construction and application of
  sparse grids \cite{bungartz2004sparse,chkifa2014high}, which attempt
  to address the problem of dimensional scaling. The sparse grid point
  sets combined with specific approximation functions allow for the
  construction of interpolation error bounds as well as bounds on
  total error for regressors by considering projections onto
  orthogonal polynomials
  \cite{migliorati2013approximation,nobile2016convergence}. However,
  sparse grids are not studied here because they have exponential size
  in dimension making them intractable for more than tens of
  dimensions and are primarily used for problems that allow data
  points to be chosen.}{} This work extends some known results
regarding the secant method \cite{dennis1996numerical} to construct
\changed{a uniform error bound for interpolants over arbitrary
  scattered data in any dimension}{an interpolation error bound for
  problems of arbitrary dimension}. These error bounds
\changed{evidence that interpolants}{} are useful, considering
\changed{similar uniform bounds cannot be constructed}{the same cannot
  be said} for \changed{regressors}{regression algorithms} in
general. The maximum complexity of an interpolant is bounded by the
amount of data available, while the maximum complexity of a regressor
is bounded by both the amount of data and the chosen parametric
form. For this reason, generic uniform bounds are largely unobtainable
for regression techniques on arbitrary approximation problems, even
when the approximation domain is bounded.

@@

\subsection{Multilayer Perceptron Regressor}
\label{sec:mlp}
The neural network is a well studied and widely used method for both
regression and classification tasks
\cite{hornik1989multilayer,rumelhart1988learning}. When using the
rectified linear unit (ReLU) activation function
\cite{dahl2013improving} and fitting the model with stochastic
gradient descent (SGD) or BFGS minimization techniques
\cite{goh2017why,moller1993scaled,robbins1951stochastic}, the model
built by a multilayer perceptron uses layers $l : \mathbb{R}^{i}
\rightarrow \mathbb{R}^{j}$ defined by
 $$ l(u) = \bigl( u^t W_l + c_l \bigr)_+ ,$$
where $c_l \in \mathbb{R}^j$ and $W_l$ is the $i \times j$ weight
matrix for layer $l$. In this form, the multilayer perceptron (MLP)
produces a piecewise linear model of the input data. The computational
complexity of training a multilayer perceptron is $\mathcal{O}(n d
m)$, where $m$ is determined by the sizes of the layers of the network
and the stopping criterion of the minimization used for finding
weights. This paper uses an MLP built from Keras and Tensorflow to
perform regression \cite{tensorflow2015-whitepaper,chollet2015keras}
with ten hidden layers each having thirty two nodes (a total of
approximately ten thousand parameters), ReLU activation, and
\changed{one hundred}{five} thousand \changed{steps}{epochs} of SGD
for training. \changed{It should be noted that the choice of neural
  network architecture affects approximation performance, but no
  architecture tuning is performed here}{}.

@@

The Delaunay triangulation is a well-studied geometric technique for
producing an interpolant \cite{lee1980two}. The Delaunay triangulation
of a set of \changed{}{data} points into simplices is such that there
are no \changed{}{data} points inside the sphere defined by the
vertices of each simplex. For a $d$-simplex S with vertices $x^{(0)}$,
$x^{(1)}$, $\ldots$, $x^{(d)}$, and function values
$f\bigl(x^{(i)}\bigr)$, $i=0$, $\ldots$, $d$, $y \in S$ is a unique
convex combination of the vertices,
 $$ y = \sum_{i=0}^{d} w_i x^{(i)}, \quad \sum_{i=0}^{d} w_i = 1, \quad w_i \geq 0, \quad i=0,\ldots,d, $$
and the Delaunay interpolant $\hat f(y)$ at $y$ is given by
 $$ \hat f(y) = \sum_{i=0}^{d} w_i f\bigl(x^{(i)}\bigr). $$ The
computational complexity of Delaunay interpolation (for the
implementation used \cite{chang2018polynomial}) is $\mathcal{O}(k n
d^3)$. Given pathological data the entire triangulation could be
computed with $k = n^{\lceil d / 2 \rceil}$, but the average case yields
$k = d \log d$ and is capable of scaling reasonably to $d \leq 50$. In
the present application, a Delaunay simplex $S$ containing $y$ is
found, then the $d+1$ vertices (points in $X$) of $S$ are used to
assign weights to each vertex and produce the predicted function value
for point $y$.

@@

The modified Shepard method used here (ShepMod) generates a continuous
approximation based on Euclidean distance and resembles a nearest
neighbor interpolant \cite{cover1967nearest}. This model is a type of
\textit{metric interpolation}, also called a Shepard method
\cite{gordon1978shepard,shepard1968two}. The interpolant has the form
 $$ p(x) = \frac{\sum\limits_{k=1}^{n}W_k(x)f\bigl(x^{(k)}\bigr)}
     {\sum\limits_{k=1}^{n}W_k(x)} ,$$ where
     \changed{$p\bigl(x^{(k)}\bigr) = f\bigl(x^{(k)}\bigr)$, $W_k(x) =
       \bigl(\bigl(r_k - \bigl\|x - x^{(k)}\bigr\|_2\bigr)_+ \big/
       \bigl( r_k \bigl\|x - x^{(k)}\bigr\|_2 \bigr)\bigr)^2$ for $x
       \neq x^{(k)}$,}{} and $r_k \in \mathbb{R}$ is the smallest
     radius such that at least $d+1$ other points $x^{(j)}$, $j \not =
     k$, are inside the closed Euclidean ball of radius $r_k$ about
     $x^{(k)}$. \changed{The interpolant $p(x)$ is continuous because
       the singularities at the $x^{(k)}$ are removable.}{} The
     computational complexity of ShepMod is $\mathcal{O}(n^2d)$. This
     paper uses a Fortran 95 implementation of ShepMod derived from
     SHEPPACK \cite{thacker2010algorithm}.

\changed{
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth,]{example_shepard_radius.pdf}
  \hspace{.2cm}
  \includegraphics[width=0.4\textwidth,]{example_shepard_weight.pdf}
  \caption{\color{red} On the left above is a depiction of the radius
    of influence for three chosen points of a collection in two
    dimensions using the modified Shepard criteria. On the right a
    third axis shows the relative weight for the center most
    interpolation point $x^{(i)}$ with the solid line representing its
    radius of influence, where $W_i(x)$ is $0$ for $\|x-x^{(i)}\|_2
    \geq r_i$ and $W_i(x) / \sum_{k=1}^n W_k(x) \to 1$ as $x \to
    x^{(i)}$.
  \vspace{-.1cm}}
  \label{fig:shepard}
\end{figure}
}{}

@@

  Suppose $g'(0) > \g/2$. Then $|g'(0)-g'(t)| \le \g t \implies g'(0)
  - \g t \le g'(t)$, and the line $w=g'(0)-\g t$ intersects the
  $t$-axis at $\tilde t=g'(0)/\g > 1/2$. $\tilde t<1$ necessarily,
  since by Rolle's Theorem there exists $0<z<1$ such that $g'(z)=0$
  and so $g'(0) - \g z \le g'(z) = 0$. Now \changed{by integrating}{}
  $$\frac{g'(0)^2}{2\g} = \frac{g'(0)\tilde t}{2} = \int_0^{\tilde t}
  (g'(0)-\g t)dt \le \int_0^{\tilde t} g'(t)dt = g(\tilde t),$$ and
  \changed{using $g'(0) > \g / 2$}{}
  \begin{align*}
    \frac{-g'(0)^2}{2\g} < g'(0) - \frac{g'(0)^2}{2\g} - \g/2 =
    \frac{(1-\tilde t)(g'(0)-\g)}{2} &= \int_{\tilde t}^1 (g'(0)-\g
    t)dt \\ &\le \int_{\tilde t}^1 g'(t)dt = -g(\tilde t),
  \end{align*}
  \noindent a contradiction for the value of $g(\tilde t)$. A similar
  contradiction arises for $g'(0)<-\g/2$. Therefore $|g'(0)| \le
  \g/2$. In terms of $f$,
  $$\bigl|\langle \nabla f(x_0)-c,x_j-x_0\rangle\bigr| = |g'(0)| \le
  \gamma \|x_j-x_0\|_2^2/2 \le \gamma k^2 / 2,$$ which holds for all
  $1\le j\le d$.  Finally, using {\it Lemma \ref{lemma:2}},
  $$\|\nabla f(x_0)-c\|_2^2 \le \frac{d}{\sigma_d^2}(\gamma k^2/2)^2
  \implies \|\nabla f(x_0)-c\|_2 \le \frac{\sqrt{d} \, \gamma
    k^2}{2\sigma_d}.$$ \qed
\end{proofdot}

@@

In addition to the figures displaying approximation results for each
data set, \changed{scatter plots of predicted versus actual values
  and}{} tables of accompanying numerical results \changed{including
  timings and quartiles of prediction error}{} are located in the
Appendix (Section \ref{sec:appendix}). All of the test data sets
capture underlying functions that are almost certainly stochastic. As
described in Section \ref{sec:introduction}, regression techniques
appear most appropriate for these problems. However, typically data
grows exponentially more sparse with increasing dimension. Given that
sparse data regressors tend towards interpolation and, as demonstrated
in Section \ref{sec:analytic}, interpolants produce similar (if not
identical) results, it is presumed that interpolants are equally
viable approximation techniques on these problems.

@@

\bibliographystyle{spmpsci}      % mathematics and physical sciences
\bibliography{tchlux-NA-interpolation}

@@

\cbend{}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth,]{scatter_forestfires.pdf}
  \hspace{.6cm}
  \includegraphics[width=0.45\textwidth,]{scatter_parkinsons.pdf}
  \vspace{.6cm}

  \includegraphics[width=0.45\textwidth,]{scatter_weather.pdf}
  \hspace{.6cm}
  \includegraphics[width=0.45\textwidth,]{scatter_creditcard.pdf}
  \vspace{.3cm}

  \caption{\color{red} Scatter plots for predicted versus actual
    values for the top three models on each of the four real valued
    approximation problems. Top left is forest fire data, top right is
    Parkinson's data, bottom left is rainfall data, and bottom right
    is credit card transaction data. The blue circles correspond to
    the best model, green diamonds to the second best, and red crosses
    to the third best model for each data set. There are a large of
    number of 0-valued entries in the forest fire and rainfall data
    sets that are not included in the visuals making the true ranking
    of the models appear to disagree with the observed outcomes.}
  \label{fig:scatter_plots}
\end{figure}
\cbend{}

@@

\begin{table}[H]
  \centering
  \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
    \hline
    Algorithm & \multicolumn{2}{c}{Min}       & \multicolumn{2}{c}{$25^{th}$} & \multicolumn{2}{c}{$50^{th}$} & \multicolumn{2}{c}{$75^{th}$} & \multicolumn{2}{c}{Max}\\
    \hline
    MARS      & $0$&$00984$                   & $3$&$11$                      & $7$&$01$                      & $\mathit{11}$&$\mathit{7}$    & $1090$&$0$\\
    SVR       & $0$&$0118$                    & $\mathbf{0}$&$\mathbf{615}$   & $\mathbf{0}$&$\mathbf{931}$   & $\mathbf{5}$&$\mathbf{89}$    & $1090$&$0$\\
    MLP       & $0$&$0426$                    & $2$&$63$                      & $5$&$27$                      & $14$&$0$                      & $1090$&$0$\\
    Delaunay  & $\mathbf{0}$&$\mathbf{0}$     & $1$&$98$                      & $5$&$37$                      & $13$&$1$                      & $\mathit{1080}$&$\mathit{0}$\\
    ShepMod   & $\mathbf{0}$&$\mathbf{0}$     & $1$&$93$                      & $6$&$27$                      & $16$&$0$                      & $1090$&$0$\\
    LSHEP     & $0$&$04$                      & $2$&$17$                      & $8$&$87$                      & $19$&$1$                      & $\mathbf{1070}$&$\mathbf{0}$\\
    Voronoi   & $\mathit{0}$&$\mathit{00982}$ & $3$&$65$                      & $7$&$56$                      & $15$&$6$                      & $1090$&$0$\\
    BoxSpline & $\mathbf{0}$&$\mathbf{0}$     & $\mathit{1}$&$\mathit{27}$    & $\mathit{4}$&$\mathit{61}$    & $12$&$6$                      & $1090$&$0$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-forest-fire}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum absolute errors respectively. The
    minimum of each column is boldface, while the second lowest value
    is italicized. All values are rounded to three significant
    digits.\vspace{.2cm}}
  \label{table:error-forest-fire}

  \begin{tabular}{|c|r@{.}l| c |r@{.}l|r@{.}l|r@{.}l|}
    \cline{1-3}\cline{5-10}
    Algorithm & \multicolumn{2}{c|}{\% Best}   &  & \multicolumn{2}{c|}{Fit/Prep. Time (s)} & \multicolumn{2}{c|}{App. Time (s)} & \multicolumn{2}{c|}{Total App. Time (s)}\\
    \cline{1-3}\cline{5-10}
    MARS      & \quad$7$&$3$                   &  & \quad\quad$29$&$1$                      & \quad$0$&$00137$                   & \quad\quad\quad$0$&$0686$\\
    SVR       & $\mathbf{78}$&$\mathbf{0}$     &  & $\mathbf{0}$&$\mathbf{00584}$           & $\mathit{0}$&$\mathit{0000620}$    & $\mathit{0}$&$\mathit{00310}$\\
    MLP       & $0$&$0$                        &  & $32$&$8$                                & $0$&$000871$                       & $0$&$0436$\\
    Delaunay  & $0$&$2$                        &  & $0$&$0151$                              & $0$&$0234$                         & $1$&$18$\\
    ShepMod   & $2$&$0$                        &  & $\mathit{0}$&$\mathit{00634}$           & $0$&$0000644$                      & $0$&$00322$\\
    LSHEP     & $5$&$1$                        &  & $0$&$0275$                              & $\mathbf{0}$&$\mathbf{0000463}$    & $\mathbf{0}$&$\mathbf{00231}$\\
    Voronoi   & $0$&$0$                        &  & $0$&$0152$                              & $0$&$000396$                       & $0$&$0198$\\
    BoxSpline & $\mathit{9}$&$\mathit{7}$      &  & $0$&$00724$                             & $0$&$0000978$                      & $0$&$00489$\\
    \cline{1-3}\cline{5-10}
  \end{tabular}
  \caption{The left above shows how often each algorithm had the
    lowest absolute error approximating forest fire data in Table
    \ref{table:error-forest-fire}. On the right columns are median fit
    time of 454 points, median time for one approximation, and median
    time approximating 50 points.\vspace{-1cm}}
  \label{table:best-forest-fire}
\end{table}

\begin{table}[H]
  \centering
  \begin{tabular}{c|r@{.}l|r@{.}l|r@{.}l|r@{.}l|r@{.}l}
    \hline
    Algorithm & \multicolumn{2}{c}{Min}                   & \multicolumn{2}{c}{$25^{th}$} & \multicolumn{2}{c}{$50^{th}$} & \multicolumn{2}{c}{$75^{th}$} & \multicolumn{2}{c}{Max}\\
    \hline
    MARS      & $0$&$00948$                               & $9$&$98$                      & $20$&$4$                      & $32$&$9$                      & $863$&$0$\\
    SVR       & $0$&$00233$                               & $3$&$15$                      & $7$&$21$                      & $11$&$3$                      & $\mathbf{28}$&$\mathbf{6}$\\
    MLP       & $0$&$0000239$                             & $\mathit{0}$&$\mathit{533}$   & $\mathit{1}$&$\mathit{25}$    & $\mathbf{2}$&$\mathbf{84}$    & $39$&$3$\\
    Delaunay  & $\mathit{3}$&$\mathit{72 \times 10^{-12}}$ & $1$&$2$                       & $3$&$5$                       & $7$&$67$                      & $30$&$7$\\
    ShepMod   & $\mathbf{0}$&$\mathbf{0}$                 & $\mathbf{0}$&$\mathbf{255}$   & $\mathbf{0}$&$\mathbf{908}$   & $\mathit{3}$&$\mathit{43}$    & $34$&$5$\\
    LSHEP     & $0$&$00254$                               & $2$&$93$                      & $7$&$16$                      & $13$&$1$                      & $\mathit{29}$&$\mathit{0}$\\
    Voronoi   & $\mathbf{0}$&$\mathbf{0}$                 & $1$&$29$                      & $3$&$52$                      & $6$&$87$                      & $30$&$1$\\
    BoxSpline & $0$&$006$                                 & $4$&$3$                       & $8$&$91$                      & $15$&$1$                      & $45$&$3$\\
    \hline
  \end{tabular}
  \caption{This numerical data accompanies the visual provided in
    Figure \ref{fig:error-parkinsons}. The columns of absolute error
    percentiles correspond to the minimum, first quartile, median,
    third quartile, and maximum absolute errors respectively. The
    minimum of each column is boldface, while the second lowest value
    is italicized. All values are rounded to three significant
    digits.\vspace{.2cm}}
  \label{table:error-parkinsons}
